Final Project Code

### 2022-12-12 ClusteringModel.Rmd
Final Code

imdb_hc_full2 <- imdb_clean %>%
  select(Budget, Runtime)

@@

two_imdb_clean <- imdb_clean %>%
    mutate(
        hclust_num = factor(cutree(imdb_full_avg2, k = 2)) # Cut into 6 clusters (k)
    )
```

@@

ggplot(two_imdb_clean, aes(x = hclust_num, fill = New_Genre)) +
    geom_bar(position = "fill") +
    labs(x = "Cluster") + 
    theme_classic()
```

>>>>

imdb_hc_full2 <- imdb_clean %>%
  select(Budget, Revenue)

@@

two_imdb_clean <- imdb_clean %>%
    mutate(
        hclust_num = factor(cutree(imdb_full_avg2, k = 3)) # Cut into 6 clusters (k)
    )
```

@@

ggplot(two_imdb_clean, aes(x = hclust_num, fill = New_Genre)) +
    geom_bar(position = "fill") +
    labs(x = "Cluster") + 
    theme_classic()

one_imdb_clean %>%
  count(hclust_num)

two_imdb_clean %>%
  count(hclust_num)
```

### 2022-12-12 Project_Final_Code.Rmd
Final Code

```{r}
full_lm_model <- fit(full_lm_wf, data = imdb_clean)

@@

```{r}
full_lm_modelcv <- fit_resamples(full_lm_wf, resamples = data_cv10, metrics = metric_set(rmse, rsq, mae))

@@

# Tune Model (trying a variety of values of Lambda penalty)

@@

```{r}
# Fit Final Model

@@

```{r}
lasso_fit %>%
  collect_metrics()
```

@@

### Visualization for Non-Linear Functions

@@

gam1_output %>%
    rsq(truth = Gross, estimate = .pred)
```

@@

```{r}
spline_rec <- recipe(Gross ~ Runtime + IMDB_Rating + Meta_score + 
                   No_of_Votes + Genre_1, data = imdb_clean) %>%
    step_nzv(all_predictors()) %>% 
    step_normalize(all_numeric_predictors()) %>% 
    step_dummy(all_nominal_predictors()) %>%
    step_naomit(Gross) %>%
    step_ns(Runtime, deg_free = 20) %>%
    step_ns(No_of_Votes, deg_free = 10) %>%
    step_ns(IMDB_Rating, deg_free = 10)

@@

@@

```{r}
cv_output_spline2 %>% collect_metrics()
```

@@

```{r}
gam1_output %>%
    rmse(truth = Gross, estimate = .pred)

@@

The GAM created using TidyModels performs better than the recipe and GAMs. Likely has to do with the degrees of freedom of the splines.

@@

# Visualize Residuals
gam1_output <- new_mod %>%
    predict(new_data = imdb_clean) %>%
    bind_cols(imdb_clean) %>%
    mutate(resid = Gross - .pred)

@@

There does not appear to be any significant bias after analyzing the residuals.

@@

```{r}
full_lm_modelcv %>% collect_metrics() #OLS
```

```{r}
lasso_fit %>% collect_metrics() #LASSO
```

```{r}
gam1_output %>%
    rmse(truth = Gross, estimate = .pred) #GAM
```

The GAM with Splines obtained using TidyModels performs the best with the lowest RMSE value (1.3 compared to 1.4 for the others).

@@

```{r}
imdb_class <- imdb_top_1000 %>%
  cSplit("Genre", sep = ",", direction = "wide") %>%
  mutate(Gross = Revenue-Budget)

@@

imdb_clean_class <- imdb_class %>%
  mutate(success_ratio = Revenue/Budget) %>%
  mutate(flop = as.factor(ifelse(success_ratio > 2, 'FALSE', 'TRUE'))) %>%
  drop_na(flop, No_of_Votes,Runtime, IMDB_Rating,Meta_score,Genre_1)
```

@@

# Create Workflows
 data_wf_mtry3 <- workflow() %>%
  add_model(rf_spec %>% set_args(mtry = 3)) %>%
  add_recipe(data_rec)

@@

```{r}
# Fit Models
set.seed(123)
data_fit_mtry3 <- fit(data_wf_mtry3, data = imdb_clean_class)

@@

```{r}
# Custom Function to get OOB predictions, true observed outcomes and add a user-provided model label
rf_OOB_output <- function(fit_model, model_label, truth){
    tibble(
          .pred_class = fit_model %>% extract_fit_engine() %>% pluck('predictions'), #OOB predictions
          flop = truth,
          model = model_label
      )
}
```

@@

@@

data_rf_OOB_output %>% 
  group_by(model) %>%
  accuracy(truth = flop, estimate = .pred_class) %>%
  mutate(mtry = as.numeric(stringr::str_replace(model,'mtry',''))) %>%
  ggplot(aes(x = mtry, y = .estimate )) + 
  geom_point() +
  geom_line() +
  theme_classic()
```

### Evaluating Selected Model - Confusion Matrix

```{r}
rf_OOB_output(data_fit_mtry4,4, imdb_clean_class %>% pull(flop)) %>%
    conf_mat(truth = flop, estimate= .pred_class)
```

@@

# Fit Model
log_fit <- fit(log_wf, data = imdb_clean_class)

@@

```{r}
# Creation of CV Folds
data_cv10_class <- vfold_cv(imdb_clean_class, v = 10)
```

```{r}
log_modelcv <- fit_resamples(log_wf, resamples = data_cv10_class, metrics = metric_set(accuracy,sens,yardstick::spec))

@@

```{r}
# Thresholds in terms of reference level
threshold_output <- final_output %>%
    threshold_perf(truth = flop, estimate = .pred_FALSE, thresholds = seq(0,1,by=.01))

# J-index v. Threshold for no flop
threshold_output %>%
    filter(.metric == 'j_index') %>%
    ggplot(aes(x = .threshold, y = .estimate)) +
    geom_line() +
    labs(y = 'J-index', x = 'threshold') +
    theme_classic()
```

@@

@@

@@

@@

We manually performed the hard predictions with threshold = .78.

@@

### Create Clusters

```{r}
imdb_sub <- imdb_clean %>%
    select(Budget, Runtime)

@@

```{r}
# Visualize the cluster assignments on the original scatterplot
imdb_clean %>%
  ggplot(aes(x = Budget, y = Runtime, color = kclust_8)) +
    geom_point() + theme_classic()
```

@@

## Hierarchial Clustering

>>>>

```{r}
# All predictors included

full_lm_model <- fit(full_lm_wf, data = imdb_clean)

@@

```{r}
# Cross Validation Performed (10 folds)

full_lm_modelcv <- fit_resamples(full_lm_wf, resamples = data_cv10, metrics = metric_set(rmse, rsq, mae))

@@

```{r}
# Tune Model (to select best Lambda penalty) -- via Cross Validation

@@

```{r}
# Fit Final Model -- Subset of Variables

@@

```{r}
# Cross Validation Performed (10 folds)

lasso_fit %>%
  collect_metrics()
```

@@

### Visualizations for Non-Linear Functions

@@

gam1_output %>%
    rsq(truth = Gross, estimate = .pred)

# Note: Unable to obtain std error for eval metrics bc can't perform cross validation w/ TidyModels
```

@@

```{r}
spline_rec <- recipe(Gross ~ Runtime + IMDB_Rating + Meta_score + 
                   No_of_Votes + Genre_1, data = imdb_clean) %>%
    step_nzv(all_predictors()) %>% 
    step_normalize(all_numeric_predictors()) %>% 
    step_dummy(all_nominal_predictors()) %>%
    step_naomit(Gross) %>%
    step_ns(Runtime, deg_free = 20) %>%
    step_ns(No_of_Votes, deg_free = 10) %>%
    step_ns(IMDB_Rating, deg_free = 10)
# Chose degrees of freedom to attempt to match # of knots from TidyModels
# May not be an exact match, but it's close

@@

GAM2_fit <- fit(spline_wf, data = imdb_clean)

tidy(GAM2_fit)
```

```{r}
# Cross Validation Performed (10 folds)

@@

@@

```{r}
# Compare RMSE values --> pick model with lower RMSE

@@

The GAM created using TidyModels performs better than the recipe GAM based on RMSE. Likely has to do with the degrees of freedom of the splines. However, it is worth noting that we were unable to perform cross validation on the TidyModels GAM.

@@

```{r}
# Visualize Residuals for Final GAM (TidyModels GAM)

@@

There does not appear to be any significant bias after analyzing the residuals. The trendline for Residuals vs. Runtime trends up at the end points since there are only a few cases with very short and very long runtimes.

@@

```{r}
# Eval Metrics from CV

full_lm_modelcv %>% collect_metrics() # OLS
```

```{r}
# Eval Metrics from CV

lasso_fit %>% collect_metrics() # LASSO
```

```{r}
# Eval Metrics from CV

gam1_output %>%
    rmse(truth = Gross, estimate = .pred) # GAM
```

The GAM with Splines obtained using TidyModels performs the best with the lowest RMSE value (approx. 1.3 compared to approx. 1.4 for the others).

@@

```{r}
# Remove log transformation from Gross

imdb_class <- imdb_top_1000 %>%
  cSplit("Genre", sep = ",", direction = "wide") %>%
  mutate(Gross = Revenue-Budget)

@@

imdb_clean_class <- imdb_class %>%
  mutate(success_ratio = Revenue/Budget) %>%
  mutate(flop = as.factor(ifelse(success_ratio > 2, 'FALSE', 'TRUE'))) %>%
  drop_na(flop, No_of_Votes, Runtime, IMDB_Rating, Meta_score, Genre_1)
```

@@

# Create Workflows to Test Various Values for mtry

 data_wf_mtry3 <- workflow() %>%
  add_model(rf_spec %>% set_args(mtry = 3)) %>%
  add_recipe(data_rec)

@@

```{r}
# Fit Models (for each mtry value)

set.seed(123)
data_fit_mtry3 <- fit(data_wf_mtry3, data = imdb_clean_class)

@@

```{r}
# Custom Function to get OOB predictions, true observed outcomes and add a user-provided model label

rf_OOB_output <- function(fit_model, model_label, truth){
    tibble(
          .pred_class = fit_model %>% extract_fit_engine() %>% pluck('predictions'), #OOB predictions
          flop = truth,
          model = model_label
      )
}
```

@@

# OOB Accuracy for Each mtry Value

@@

data_rf_OOB_output %>% 
    group_by(model) %>%
    sensitivity(truth = flop, estimate = .pred_class)
# mtry = 3 and 4 tied for best sensitivity

# Accuracy vs. mtry Plot

data_rf_OOB_output %>% 
  group_by(model) %>%
  accuracy(truth = flop, estimate = .pred_class) %>%
  mutate(mtry = as.numeric(stringr::str_replace(model,'mtry',''))) %>%
  ggplot(aes(x = mtry, y = .estimate )) + 
  geom_point() +
  geom_line() +
  theme_classic()
# Select mtry = 4 based on accuracy
```

### Evaluating Selected Model - Confusion Matrix and Evaluation Metrics

```{r}
# Confusion Matrix

rf_OOB_output(data_fit_mtry4,4, imdb_clean_class %>% pull(flop)) %>%
    conf_mat(truth = flop, estimate= .pred_class)

# Sensitivity, Specificity, and Accuracy

rf_OOB_output(data_fit_mtry4,4, imdb_clean_class %>% pull(flop)) %>%
  sensitivity(truth = flop, estimate= .pred_class)

rf_OOB_output(data_fit_mtry4,4, imdb_clean_class %>% pull(flop)) %>%
  specificity(truth = flop, estimate= .pred_class)

rf_OOB_output(data_fit_mtry4,4, imdb_clean_class %>% pull(flop)) %>%
  accuracy(truth = flop, estimate= .pred_class)
```

@@

```{r}
# Fit Model

log_fit <- fit(log_wf, data = imdb_clean_class)

@@

```{r}
# Creation of CV Folds

data_cv10_class <- vfold_cv(imdb_clean_class, v = 10)
```

```{r}
# Cross Validation Performed (10 folds)

log_modelcv <- fit_resamples(log_wf, resamples = data_cv10_class, metrics = metric_set(accuracy,sens,yardstick::spec))

@@

```{r}
# Thresholds in terms of reference level

threshold_output <- final_output %>%
    threshold_perf(truth = flop, estimate = .pred_FALSE, thresholds = seq(0,1,by=.01))

# J-index v. Threshold for no flop

threshold_output %>%
    filter(.metric == 'j_index') %>%
    ggplot(aes(x = .threshold, y = .estimate)) +
    geom_line() +
    labs(y = 'J-index', x = 'threshold') +
    theme_classic()
```

@@

# Compare Eval Metrics

@@

# Compare Confusion Matrices

@@

We chose the threshold of 0.78 for our final model because we are prioritizing sensitivity over accuracy and specificity.

@@

We manually performed the hard predictions with threshold = .78. Since .pred_FALSE = 0.56 which is lower than our threshold of 0.78, we would predict this example movie would flop (flop = TRUE).

@@

### Feature Selection

```{r}
imdb_sub <- imdb_clean %>%
    select(Budget, Gross)

@@

```{r}
# Visualize the cluster assignments on the original scatterplot
imdb_clean %>%
  ggplot(aes(x = Budget, y = Gross, color = kclust_8)) +
    geom_point() + theme_classic()
```

@@

## Hierarchical Clustering
